COFFEE: an Optimizing Compiler\\for Finite Element Local Assembly

### 2014-04-07 paper/bare_conf.tex
WIP on performance section

Experiments were run on two Intel machines, a Sandy Bridge (I7-2600 CPU, running at 3.4GHz, 32KB L1 cache and 256KB L2 cache) and the Phi. The \texttt{icc 13.1} compiler was used, with optimization level \texttt{-O2} and with auto-vectorization enabled (\texttt{-xAVX} on the Sandy Bridge, and \texttt{TODO} on the Phi). Other optimization levels performed, in general, slightly worse than \texttt{-O2}. Our code transformations were evaluated in three real-world problems based on the following PDEs:
\begin{itemize}
\item Helmholtz
\item Advection-Diffusion
\item Burgers
\end{itemize}
The code was written in UFL and then executed over real unstructured meshes through Firedrake. The Helmholtz code has already been shown in Figure~\ref{code:helmholtz}. For Advection-Diffusion, the ``Diffusion'' equation, which uses the same differential operators as Helmholtz, is considered. In the Diffusion kernel, the main differences with respect to Helmholtz are the absence of the $Y$ array and the presence of a few more constants for computing the stiffness matrix $A$. Burgers is a time-dependent problem, i.e. the assembly is recalculated every time step based on the result of previous iterations. It employs differential operators different from those of Helmholtz, which has a major impact on the generated assembly code (Figure~\ref{code:burgers}), where a larger number of basis function matrices ($X1$, $X2$, ...) and constants ($F0$, $F1$, ..., $K0$, $K1$, ...) are accessed.

These problems were studied varying both the shape of mesh elements and the polynomial order $p$ of the method. Intuitively, the bigger the element shape and $p$, the larger is the iteration space. Triangles (2D), tetrahedron (3D), and prisms (3D) were tested as element shape. For instance, in the case of Helmholtz with $p=1$, the size of the $j$ and $k$ loops for the three kind of elements is, respectively, $3$, $4$, and $6$. Moving from 2D to 3D has the effect of increasing the number of basis function arrays, since, conceptually, the behaviour of the equation has now to be approximated also along a third axis. On the other hand, the polynomial order affects only the problem size (the three loops $i$, $j$, and $k$, and, as a consequence, the size of $X$ and $Y$ arrays). A range of polynomial orders, from $p=1$ to $p=4$, were tested; higher polynomial orders are excluded from the study because of current Firedrake limitations. In such a large space of problems, the size of the stiffness matrix rarely exceeds 30$\times$30, with a peak of 105$\times$105 in Burgers with prisms and $p=4$.

>>>>

% TODO Xeon Phi flags
% FIXME: cite UFL

Experiments were run on two Intel architectures, a Sandy Bridge (I7-2600 CPU, running at 3.4GHz, 32KB L1 cache and 256KB L2 cache) and the Xeon Phi. The \texttt{icc 13.1} compiler was used, with optimization level \texttt{-O2} and with auto-vectorization enabled (\texttt{-xAVX} on the Sandy Bridge, and \texttt{TODO} on the Phi). Other optimization levels performed, in general, slightly worse than \texttt{-O2}. Our code transformations were evaluated in three real-world problems based on the following PDEs:
\begin{itemize}
\item Helmholtz
\item Advection-Diffusion
\item Burgers
\end{itemize}
The code was written in UFL and then executed over real unstructured meshes through Firedrake. The Helmholtz code has already been shown in Figure~\ref{code:helmholtz}. For Advection-Diffusion, the ``Diffusion'' equation, which uses the same differential operators as Helmholtz, is considered. In the Diffusion kernel, the main differences with respect to Helmholtz are the absence of the $Y$ array and the presence of a few more constants for computing the element matrix $A$. Burgers is a time-dependent problem, which means the assembly is recalculated every time step based on the result of previous iterations. It employs differential operators different from those of Helmholtz, which has a major impact on the generated assembly code shown in Figure~\ref{code:burgers}, where a larger number of basis function matrices ($X1$, $X2$, ...) and constants ($F0$, $F1$, ..., $K0$, $K1$, ...) are accessed.

These problems were studied varying both the shape of mesh elements and the polynomial order $p$ of the method. Intuitively, the bigger the element shape and $p$, the larger is the iteration space. Triangles (2D), tetrahedron (3D), and prisms (3D) were tested as element shape. For instance, in the case of Helmholtz with $p=1$, the size of the $j$ and $k$ loops for the three kind of elements is, respectively, $3$, $4$, and $6$. Moving from 2D to 3D has the effect of increasing the number of basis function arrays, since, conceptually, the behavior of the equation has now to be approximated also along a third axis. On the other hand, the polynomial order affects only the problem size (the three loops $i$, $j$, and $k$, and, as a consequence, the size of $X$ and $Y$ arrays). A range of polynomial orders, from $p=1$ to $p=4$, were tested; higher polynomial orders are excluded from the study because of current Firedrake limitations. In such a large space of problems, the size of the element matrix rarely exceeds 30$\times$30, with a peak of 105$\times$105 in Burgers with prisms and $p=4$.

% FIXME: which limitations?
